---
title: "News"
lightbox: true
execute:
  echo: false
  message: false
  warning: false
---


```{r echo = F, warning=F, message=F}
library(tidyverse)

df_v1 <- readxl::read_xlsx('files/Native bees of BC - PM.xlsx')
df_v2 <- readxl::read_xlsx('files/Native bees of BC - PM_2nd edition.xlsx') %>%
  filter(`Bee book list` == 'Y')
df_v3 <- readxl::read_xlsx("files/Native bees of BC - PM_3rd edition.xlsx") %>%
  filter(`Bee book list` == 'Y')

gbif_v1 <- read_delim("files/0037810-231120084113126.csv", delim = "\t") %>%
  mutate(eventDate = str_split(eventDate, "/", simplify = TRUE)[1],
         eventDate = parse_date_time(eventDate, orders = c("ymd", "ymd HM", 'ymd HMS')),
         date = ymd(date(eventDate)), Year = year(date),
         Month = lubridate::month(date, label=T),
         lat=decimalLatitude, lon=decimalLongitude) %>%
  drop_na(date, lat, lon, genus, occurrenceID) %>%
  separate_wider_delim(species, ' ', names=c('Genus', 'Species'), too_many = 'merge') %>%
  mutate(Species=ifelse(Species=='nr', NA, Species)) %>%
  drop_na(Species)


gbif_v2 <- read_delim("files/0014018-240906103802322.csv", delim = "\t") %>%
  mutate(eventDate = str_split(eventDate, "/", simplify = TRUE)[1],
         eventDate = parse_date_time(eventDate, orders = c("ymd", "ymd HM", 'ymd HMS')),
         date = ymd(date(eventDate)), Year = year(date),
         Month = lubridate::month(date, label=T),
         lat=decimalLatitude, lon=decimalLongitude) %>%
  drop_na(date, lat, lon, genus, occurrenceID) %>%
  separate_wider_delim(species, ' ', names=c('Genus', 'Species'), too_many = 'merge') %>%
  mutate(Species=ifelse(Species=='nr', NA, Species)) %>%
  drop_na(Species)

gbif_v3 <- gbif <- read_delim("files/0050043-251120083545085.csv", delim = "\t") %>%
  separate_wider_delim(eventDate, delim='/', names = c('eventDate', NA), too_few = "align_start") %>%
  mutate(eventDate = parse_date_time(eventDate, orders = c("ymd", "ymd HM", 'ymd HMS', "ym", "y")),
         date = ymd(date(eventDate)), Year = year(date),
         Month = lubridate::month(date, label=T),
         lat=decimalLatitude, lon=decimalLongitude) %>%
  drop_na(date, lat, lon, occurrenceID) %>%
  separate_wider_delim(species, ' ', names=c('Genus', 'Species'), too_many = 'merge', too_few = 'align_start') %>%
  mutate(Species=ifelse(Species=='nr', NA, Species)) 


df_v1_list <- unique(paste(df_v1$Family, df_v1$Genus, df_v1$Species))
df_v2_list <- unique(paste(df_v2$Family, df_v2$Genus, df_v2$Species))
df_v3_list <- unique(paste(df_v3$Family, df_v3$Genus, df_v3$Species))
gbif_v1_list <- unique(paste(gbif_v1$family, gbif_v1$Genus, gbif_v1$Species))
gbif_v2_list <- unique(paste(gbif_v2$family, gbif_v2$Genus, gbif_v2$Species))
gbif_v3_list <- unique(paste(gbif_v3$family, gbif_v3$Genus, gbif_v3$Species))
```


## 3rd Edition (December 2025)

More new bees for 2025! ðŸ

::: {.callout-note icon=false}
## <i class="fa-solid fa-award"></i>  Bee Book recognized

Our efforts have been recognized by the [Native Bee Society of BC](https://www.bcnativebees.org) with the 2025 certificate of appreciation for data management ðŸŽ‰ thanks all!
:::

* Bee book has been used for the [City of North Vancouver annual report](https://www.bcnativebees.org/projects/cnv-bcba24) by the [Native Bee Society of BC](https://www.bcnativebees.org/)
* Our melittologist friends are using the Bee Book to decide where to look for bees to document
* We used the Bee Book to prepare the [Bee Atlas annual summary](https://www.bcnativebees.org/projects/2024-25-bcba-annual-report)
* New updates to introduction and general info on the Bee Book site
* 3rd edition data release

```{r echo = F, warning=F, message=F}
#| layout-ncol: 2

summary <- tibble(
  version=c(1,1,2,2,3,3), 
  date=dmy(c('01/12/23','01/12/23', '12/09/24', '12/09/24', '12/12/25', '12/12/25')), 
  ref = c("Bee Gallery", 'GBIF', "Bee Gallery", 'GBIF', "Bee Gallery", 'GBIF'),
  new = c(
    0,0,
    length(setdiff(df_v2_list, df_v1_list)), length(setdiff(gbif_v2_list, gbif_v1_list)), 
    length(setdiff(df_v3_list, df_v2_list)), length(setdiff(gbif_v3_list, gbif_v2_list))
  ),
  old = c(
    length(df_v1_list), length(gbif_v1_list),
    length(intersect(df_v1_list,df_v2_list)), length(intersect(gbif_v1_list,gbif_v2_list)), 
    length(intersect(df_v2_list,df_v3_list)), length(intersect(gbif_v2_list,gbif_v3_list))),
  removed = c(
    0,0,
    length(setdiff(df_v1_list, df_v2_list)),length(setdiff(gbif_v1_list, gbif_v2_list)),
    length(setdiff(df_v2_list, df_v3_list)),length(setdiff(gbif_v2_list, gbif_v3_list))
    )
)

summary %>% 
  rowwise() %>%
  mutate(tot = new + old) %>%
  ggplot(aes(x = ordered(version), y = tot, colour=ref)) + 
  geom_line(aes(group = ref)) + 
  geom_point() + 
  ggpubr::theme_pubr() +
  scale_colour_manual(values = c('#376D61', 'grey70'), labels = c("Bee Gallery", "GBIF")) + 
  labs(x = 'Version', y = 'Number of Species', colour ='Collection')


summary %>%
  mutate(removed = -removed) %>%
  pivot_longer(c(new, old, removed)) %>%
  mutate(tot = name != 'removed') %>%
  ggplot(aes(x = ordered(version), y = value, fill = name)) + 
  geom_col(position = position_stack()) + 
  facet_wrap(~ref, ncol=2) + 
  ggpubr::theme_pubr() +
  scale_fill_manual(values = c('#376D61', '#A0CFC4',  'grey70'), labels = c("New", "Retained", "Removed")) + 
  labs(x = 'Version', y = 'Number of species', fill = 'Since previous version')


```


### Resources

  * [GBIF doi](https://doi.org/10.15468/dl.cs2xsc)
  * [Bee book data](https://doi.org/10.5281/zenodo.18112558)
  * [Code release](https://github.com/teaswamp-creations/bee-book/releases/tag/v3.0.0)
  
### Data Completeness

Bee book gallery shows two pieces of information per species page:

1.  Bee image (curated by KQ)
2.  Map of observations in Pacific Maritime (pulled from GBIF)

Page completeness is as follows:

```{r echo=F}
upset_list = list(`Has PM map` = unique(paste(gbif_v3$family, gbif_v3$Genus, gbif_v3$Species)),
                  `In bee gallery` = unique(paste(gbif_v3$Family, gbif_v3$Genus, gbif_v3$Species)), 
                  `Has image` = unique(paste(gbif_v3$Family, gbif_v3$Genus, gbif_v3$Species)[!is.na(gbif_v3$`Image link`)])
)

UpSetR::upset(UpSetR::fromList(upset_list))
``` 




## 2nd Edition (September 2024)

New bees added! 

* Bee pages feature a summary note of historical records curated from literature (see references page)
* News page added
* References page added
* Bee badges updated to locations of interest


```{r echo = F, warning=F, message=F}
#| layout-ncol: 2

summary <- tibble(
  version=c(1,1,2,2), 
  date=dmy(c('01/12/23','01/12/23', '12/09/24', '12/09/24')), 
  ref = c("Bee Gallery", 'GBIF', "Bee Gallery", 'GBIF'),
  new = c(0,0,length(setdiff(df_v2_list, df_v1_list)), length(setdiff(gbif_v2_list, gbif_v1_list))),
  old = c(length(df_v1_list),length(gbif_v1_list),length(intersect(df_v1_list,df_v2_list)),length(intersect(gbif_v1_list,gbif_v2_list))),
  removed = c(0,0,length(setdiff(df_v1_list, df_v2_list)),length(setdiff(gbif_v1_list, gbif_v2_list)))
)

summary %>% 
  rowwise() %>%
  mutate(tot = new + old) %>%
  ggplot(aes(x = ordered(version), y = tot, colour=ref)) + 
  geom_line(aes(group = ref)) + 
  geom_point() + 
  ggpubr::theme_pubr() +
  scale_colour_manual(values = c('#376D61', 'grey70'), labels = c("Bee Gallery", "GBIF")) + 
  labs(x = 'Version', y = 'Number of Species', colour ='Collection')


summary %>%
  mutate(removed = -removed) %>%
  pivot_longer(c(new, old, removed)) %>%
  mutate(tot = name != 'removed') %>%
  ggplot(aes(x = ordered(version), y = value, fill = name)) + 
  geom_col(position = position_stack()) + 
  facet_wrap(~ref, ncol=2) + 
  ggpubr::theme_pubr() +
  scale_fill_manual(values = c('#376D61', '#A0CFC4',  'grey70'), labels = c("New", "Retained", "Removed")) + 
  labs(x = 'Version', y = 'Number of species', fill = 'Since previous version')


```


### Resources

  * [GBIF doi](https://doi.org/10.15468/dl.v6ttrj)
  * [Bee book data](https://doi.org/10.5281/zenodo.13757007)
  * [Code release](https://github.com/teaswamp-creations/bee-book/releases/tag/v2.0.0)
  
### Data Completeness

Bee book gallery shows two pieces of information per species page:

1.  Bee image (curated by KQ)
2.  Map of observations in Pacific Maritime (pulled from GBIF)

Page completeness is as follows:

```{r echo=F}
upset_list = list(`Has PM map` = unique(paste(gbif_v2$family, gbif_v2$Genus, gbif_v2$Species)),
                  `In bee gallery` = unique(paste(df_v2$Family, df_v2$Genus, df_v2$Species)), 
                  `Has image` = unique(paste(df_v2$Family, df_v2$Genus, df_v2$Species)[!is.na(df_v2$`Image link`)])
)

UpSetR::upset(UpSetR::fromList(upset_list))
``` 




## 1st Edition release (December 2023)

First published version of the bee book!

  * [GBIF doi](https://doi.org/10.15468/dl.vtr55y)
  * [Bee book data](https://doi.org/10.5281/zenodo.13755527)
  * [Code release](https://github.com/teaswamp-creations/bee-book/releases/tag/v1.0.0)
  
### Data Completeness

Bee book gallery shows two pieces of information per species page:

1.  Bee image (curated by KQ)
2.  Map of observations in Pacific Maritime (pulled from GBIF)

Page completeness is as follows:
  
```{r echo=F}
upset_list = list(`Has PM map` = unique(paste(gbif_v1$family, gbif_v1$Genus, gbif_v1$Species)),
                  `In bee gallery` = unique(paste(df_v1$Family, df_v1$Genus, df_v1$Species)), 
                  `Has image` = unique(paste(df_v1$Family, df_v1$Genus, df_v1$Species)[!is.na(df_v1$`Image link`)])
)

UpSetR::upset(UpSetR::fromList(upset_list))
``` 


Last updated: `r now()`

